---
title: "Weight_problems"
author: "Fay"
date: "2024-04-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, echo=FALSE, warning=FALSE}
# ***********************************************************
# Title: Predicting the health outcomes of infections in hybrid mice

# Purpose: This script defines all the settings and executes
#         all the code (.R, .md) to reproduce the analysis
#         of the project
#
# Authors: Fay Webster
# ID variables:
# ***********************************************************
# Part 1: Set standard settings & install packages            ----
# ***********************************************************
  # Install packages/load libraries to keep R environment stable
    # install
        # pacman for simplified bulk pkg import
        # renv for pkg consistency over time
#install.packages("pacman")
# increase maximum overlaps
options(ggrepel.max.overlaps = Inf)

library(pacman)
    ## Standard settings ----
        # seed
set.seed(13102023)

        # Use p_load to install (if not already) and load the packages
pacman::p_load(mice, stringr, gridExtra, dplyr, tidyverse, tidyr, janitor, 
               visdat, corrplot, RColorBrewer, ggplot2, VIM, limma, 
               latticeExtra, patchwork,FactoMineR, ggrepel, factoextra, 
               reshape2, sjPlot, stargazer, jtools,modelsummary, ggeffects, 
               pheatmap, ggpubr, ggridges, gt, caret, randomForest, rfUtilities,
               parasiteLoad, fitdistrplus, optimx, leaflet, magick, ggdist,
               ggbeeswarm)
    

hm <- read.csv("data/analysis/final/imputed_clean_data.csv")
Challenge <- read.csv("data/lab/final/lab_cleaned_data.csv")

 
 
 ######### Decision: Removing IL.10 from gene selection due data largely missing
 # remove il_10
 #### vectors for selecting genes for analysis
 Genes_v   <- c("IFNy", "CXCR3", "IL.6", "IL.13", #"IL.10",
                "IL1RN","CASP1", "CXCL9", "IDO1", "IRGM1", "MPO", 
                "MUC2", "MUC5AC", "MYD88", "NCR1", "PRF1", "RETNLB", "SOCS1", 
                "TICAM1", "TNF")
 
 # Select laboratory data 
# Select genes
lab <- hm %>%
    dplyr::filter(origin == "Lab")

Field <- hm %>%
    dplyr::filter(origin == "Field")
# Then, define the color for each level of infection
color_mapping <- c("E_falciformis" = "salmon", 
                   "E_ferrisi" = "forestgreen", 
                   "uninfected" = "skyblue")
```


How many mice went through primary infections?
```{r}
Challenge %>%
           filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID,
                  dpi == max_dpi) %>%
    summarise(n())
```
How many mice died in the primary infections?
```{r}
Challenge %>%
           filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID,
                  death == "primary", dpi == max_dpi) %>%
    summarise(n())
```
How many mice were in the challenge infections?
```{r}
Challenge %>%
           filter(infection == "challenge", Mouse_ID %in% lab$Mouse_ID,
                  dpi == max_dpi, death == "challenge") %>%
    summarise(n()) 

```
Something is fishy here. Do we have duplicates? Yes, I forgot to remove the spleen 
measurements. Luke used the spleen for gene expression for some mice. 
We use here the mesenterial lymphnodes. (?)
```{r}
Challenge %>%
           filter(infection == "challenge", Mouse_ID %in% lab$Mouse_ID,
                  dpi == max_dpi, death == "challenge", Position == "mLN") %>%
    summarize(n())
```


Original in article of primary infections: 
I include mice that died in primary infections, but also the mice that went on 
to be challenged.

```{r}
ggplot(Challenge %>%
           filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID) %>%
           group_by(Mouse_ID), 
       aes(x = WL_max, y = Parasite_primary, fill = Parasite_primary)) + 
    geom_density_ridges(jittered_points = TRUE, 
                        position = position_points_jitter(height = 0), 
                        scale = 0.9, alpha = 0.6, point_shape = 21, 
                        point_size = 2, point_alpha = 1) +
    geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5,
                 position = position_nudge(x = 0.2)) +
    # coord_flip() +
    theme_minimal() +
    scale_fill_manual(values = color_mapping) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.3)) +
    xlab("Maximum relative weight loss") +
    ylab("Parasite strain - primary infections")
```
What happens, if I remove the mice that were challenged. 
Here, including just the mice that died in the primary infections.
```{r}
ggplot(Challenge %>%
           filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID,
                  death == "primary") %>%
           group_by(Mouse_ID), 
       aes(x = WL_max, y = Parasite_primary, fill = Parasite_primary)) + 
    geom_density_ridges(jittered_points = TRUE, 
                        position = position_points_jitter(height = 0), 
                        scale = 0.9, alpha = 0.6, point_shape = 21, 
                        point_size = 2, point_alpha = 1) +
    geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5,
                 position = position_nudge(x = 0.2)) +
    # coord_flip() +
    theme_minimal() +
    scale_fill_manual(values = color_mapping) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.3)) +
    xlab("Maximum relative weight loss") +
    ylab("Parasite strain - primary infections")
```
Where are the uninfected mice? Did we just have infected mice? I remember 
reaching the same conclusion before. 

How many mice died on each day in the primary infections?
```{r}
Challenge %>%
           filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID,
                  death == "primary", dpi == max_dpi) %>%
    group_by(max_dpi) %>%
    summarise(n())
    
Challenge %>%
           filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID,
                  death == "primary", dpi == max_dpi) %>%
    ggplot(aes(x = max_dpi)) +
    geom_histogram()

```


What about the challenge infections?
How many mice died on each day in the primary infections?
```{r}
Challenge %>%
           filter(infection == "challenge", Mouse_ID %in% lab$Mouse_ID,
                  death == "challenge", dpi == max_dpi, Position == "mLN") %>%
    group_by(max_dpi) %>%
    summarise(n())
    
Challenge %>%
           filter(infection == "challenge", Mouse_ID %in% lab$Mouse_ID,
                  death == "challenge", dpi == max_dpi, Position == "mLN") %>%
    ggplot(aes(x = max_dpi)) +
    geom_histogram()

```

# Challenge infections remains the same


```{r}

ggplot(lab %>%
           filter(infection == "challenge") %>%
           group_by(Mouse_ID), 
       aes(x = WL_max, y = Parasite_challenge, fill = Parasite_challenge)) + 
    geom_density_ridges(jittered_points = TRUE, position = 
                            position_points_jitter(height = 0), 
                        scale = 0.9, alpha = 0.6, point_shape = 21, 
                        point_size = 2, point_alpha = 1) +
    geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5, 
                 position = position_nudge(x = 0.2)) +
    # coord_flip() +
    theme_minimal() +
    scale_fill_manual(values = color_mapping) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.3)) +
    xlab("Maximum relative weight loss") +
    ylab("Parasite strain - challenge infections") 
```

So what about the weight loss per dpi. 
Here is what I originally plotted: 
```{r}
#relative weight loss per day - primary
Challenge %>%
    filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID) %>%
    ggplot(aes(x = dpi, y = relative_weight, color = Parasite_primary, 
               fill = Parasite_primary)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.4, 
                shape = 21, stroke = 0.5, size = 3) + # Adjusted for outlines
    geom_smooth(aes(fill = Parasite_primary), 
                method = "loess", se = TRUE, alpha = 0.2) + 
    # Add smooth line with confidence intervals
    scale_color_manual(values = color_mapping) + # Apply custom color mapping
    scale_fill_manual(values = color_mapping) + 
    # Ensure fills match colors for confidence intervals
    labs(#title = "Relative Weight by Days Post Infection",
        x = "Days Post Infection (dpi)",
        y = "Relative weight, primary infections",
        color = "Infection group",
        fill = "Infection group") + # Added for consistency with the legend
    theme_minimal() + # Use a minimal theme for a cleaner look
    theme(legend.position = "right", # Adjust legend position
          plot.title = element_text(hjust = 0.5), # Center the plot title
          legend.title.align = 0.5) 

```
Should I include ONLY the 20 mice that were solely subjected to primary infections?
```{r}
#relative weight loss per day - primary
Challenge %>%
    filter(infection == "primary", Mouse_ID %in% lab$Mouse_ID, death == "primary") %>%
    ggplot(aes(x = dpi, y = relative_weight, color = Parasite_primary, 
               fill = Parasite_primary)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.4, 
                shape = 21, stroke = 0.5, size = 3) + # Adjusted for outlines
    geom_smooth(aes(fill = Parasite_primary), 
                method = "loess", se = TRUE, alpha = 0.2) + 
    # Add smooth line with confidence intervals
    scale_color_manual(values = color_mapping) + # Apply custom color mapping
    scale_fill_manual(values = color_mapping) + 
    # Ensure fills match colors for confidence intervals
    labs(#title = "Relative Weight by Days Post Infection",
        x = "Days Post Infection (dpi)",
        y = "Relative weight, primary infections",
        color = "Infection group",
        fill = "Infection group") + # Added for consistency with the legend
    theme_minimal() + # Use a minimal theme for a cleaner look
    theme(legend.position = "right", # Adjust legend position
          plot.title = element_text(hjust = 0.5), # Center the plot title
          legend.title.align = 0.5) 

```
Why do the ferrisi mice loose more weight? They probably died on day 6 naturally?
Many mice fall under the 80% relative weight. 

Back to the challenge infections: 
```{r}
#relative weight loss per day - challenge
Challenge %>%
    filter(infection == "challenge", Mouse_ID %in% lab$Mouse_ID) %>%
    ggplot(aes(x = dpi, y = relative_weight, color = Parasite_challenge, 
               fill = Parasite_challenge)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.4, 
                shape = 21, stroke = 0.5, size = 3) + # Adjusted for outlines
    geom_smooth(aes(fill = Parasite_challenge), 
                method = "loess", se = TRUE, alpha = 0.2) + 
    #Add smooth line with confidence intervals
    scale_color_manual(values = color_mapping) + # Apply custom color mapping
    scale_fill_manual(values = color_mapping) + 
    # Ensure fills match colors for confidence intervals
    labs(#title = "Relative Weight by Days Post Infection",
        x = "Days Post Infection (dpi)",
        y = "Relative weight, challenge infections",
        color = "Infection group",
        fill = "Infection group") + # Added for consistency with the legend
    theme_minimal() + # Use a minimal theme for a cleaner look
    theme(legend.position = "right", # Adjust legend position
          plot.title = element_text(hjust = 0.5), # Center the plot title
          legend.title.align = 0.5)
```
Falciformis mice falling under 80%, but only on day 8.

Mouse strains: 
```{r}
# transform mouse strain into factor
lab$mouse_strain <- as.factor(lab$mouse_strain)

lab$mouse_strain <- gsub(pattern = "_", " ", lab$mouse_strain)

#Numbers of each mouse strain
ggplot(lab, aes(x = WL_max, y = mouse_strain, fill = mouse_strain)) + 
    geom_density_ridges(jittered_points = TRUE, position = 
                            position_points_jitter(height = 0), 
                        scale = 0.9, alpha = 0.6, point_shape = 21, 
                        point_size = 2, point_alpha = 1) +
    geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5, 
                 position = position_nudge(x = 0.2)) +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 70, vjust = 0.5, hjust=0.3)) +
    xlab("Maximum relative weight loss") +
    ylab("Mouse Strain") +
    facet_grid(~infection)
```



